Handling String-Number Conversion in String Constraint Solving

### 2019-10-01 draft.tex
update

\title{Handling String-number Conversion in String Constraint Solving}
\author{}
\institute{}

@@

\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\sti}[1]{\mbox{\textbf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textbf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_n$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_s$}}
\newcommand{\cvar}{\mbox{$\mathbb{C}$}}
\newcommand{\modelof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}

@@

\todo{Why this problem is important? Need to say it is used often, e.g., read 
from text file or user input. Why this problem is challenging? It is undecidable. Some tricky cases 
using string as array index. Bounded SAT-based encoding might not work even for 
very simple constraints. e.g., $y=str2int(x) \wedge y>9999999$, if $x$ is a 
ASCII string, then XXX boolean variables are required. Or we can just propose 
an example that all major solver fails. Maybe we still say our approach is still over-approximation + under-approximation, in this case, need to think how to present the over-approximation part. Maybe no CEGAR is fine.}

In this paper, we propose an approach that handles string constraints with string-number conversion efficiently. From a theoretical point of view, this problem is already proven to be undecidable~\cite{vijay_rp}, and in practice, in our experiments, all of the state-of-the-art string constraint solvers do not support string-number conversion correctly. Some of them handles such constraints with only over-approximation of the original formula, and is known to have one-side error (see Section~\ref{section:experiment} for detailed explanation). Here we propose a strategy that uses both over and under-approximation of the string constraints. The over-approximation is for proving UNSAT and the under-approximation is for proving SAT of the string constraint. Both over and under-approximation should fail in a decidable fragment of string constraints and can be efficiently solved.

For over-approximation, there are a few existing algorithms~\cite{atva2019,z3}, our strategy does not set any restriction on which algorithm to be used. In our experiences, the UNSAT instances in all available benchmarks are not that difficult and many approaches can handle all of them almost perfectly. The battle field is in fact those SAT instances.

More specifically, instead of allowing the variable domain to be arbitrary strings from a finite alphabet, we restrict the search space of each string variable to strings that obey some predefined and parameterized pattern. By adjusting the parameters, one can easily enlarge or prune out the potential solution space. In paper, we propose to use patterns defined by \emph{symbolic flat automata}, the automata are very expressive yet allows very efficiently manipulation. For example, it avoids the alphabet 
explosion problem that the approach in~\cite{PLDI2017} suffers.

Based on the symbolic flat automata encoding, we manage to convert the 
string-number function to a constraint consisting of both 
polynomials and exponentials. To the best of our knowledge, the satisfiability 
problem of a mix of polynomials and exponentials is still open. For the 
particular case with only one number variable exponent, the satisfiability can 
be decided using the algorithm in~\cite{POPL2019}. However, the algorithm involves a quantifier elimination procedure, which is double-exponential to the length of the input formula and hence is not scalable to large instances.  \todo{check correctness of the above statement} For bounded numbers, our approach converts string-number function to a linear integer constraint.
To ease presentation, in the main text, we only consider string constraints consisting of \emph{equality constraints (a.k.a. word equation)}, \emph{length constraints}, and \emph{string-number conversion} functions. We describe the extension to support transducer and membership constraints in the appendix.

@@

A \emph{finite-state automaton} (FA) is a tuple $(Q,T,\Sigma,q_0,q_m)$, where $Q$ is the set of states, $T\subseteq Q\times \Sigma\times Q $ is the set of transition relation, $\Sigma$ is the alphabet, $q_0$ is the initial state, and $q_m$ is the final state. The semantics of FSA is defined in the standard manner. We use $L(A)$ to denote the regular language, i.e., set of accepted strings, of a FA $A$.

Given a FA $A$ over $\{a_1,a_2,\ldots,a_n\}$, the Parikh image of $L$ is the set $\{(|w|_{a_1},\ldots, |w|_{a_n}) \mid w \in  L(A) \}$. Moreover, one can construct from $A$ in linear time a Presburger formula $\phi_A(v_1, \ldots,v_n)$ such that $\phi(c_1, \ldots,c_n) \iff (c_1, \ldots,c_n)$ in the Parikh image of $L$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{String Constraints} \label{section:sc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Fix a finite alphabet $\Sigma \subseteq \mathbb{N}$, a finite set of string variables $\vars$ ranging over $\Sigma^*$, and a finite set of natural number variables $\varn$. Note that here we assume the alphabet is a finite subset of natural numbers. Essentially, we try to capture the numerical encoding of symbols (e.g., in ASCII, `A' is encoded as $65$) in computers. We write $[n,m]$ for the set of numbers (or symbols in $\Sigma$) $\{k\mid n\leq k \leq m\}$. A \emph{term} is an element in $(\vars\cup \Sigma)^*$. An \emph{interpretation} is a mapping from $\vars\cup \varn$ to $\Sigma^* \cup \mathbb{N}$. We lift 
it to terms by defining $I: (\vars\cup \Sigma)^* \rightarrow \Sigma^*$ with $I(a) =a$ for $a\in \Sigma$ and $I(t_1 \cdot t_2) = I(t_1)\cdot I(t_2)$.

The \emph{string-number conversion function} $\sti{x}$ is defined as follows. For $x\in [0,9]^+$ and $a\in [0,9]$, we have $\sti{a}=a$ and $\sti{x\cdot a} = 10\times \sti{x}+a$. For $x\notin [0,9]^+$, $\sti{x}=-1$.

For a number $n=a_n*10^n+a_{n-1}*10^{n-1}+\ldots+a_0$ for $a_i\in [0,9], i\in [0,n]$, the \emph{number-string conversion function} $\its{n}=a_ia_{i-1}\ldots a_0$~\footnote{Notice that only positive integer is supported in this semantics. This is the semantics used by most of the SMT solvers, and hence we follow it in this paper. This is not a restriction, since conversing from negative integer can still be encoded using the positive only version.}.

@@

An \emph{integer constraint} $\phi$ is a linear constraint over the variables in $\varn$, numerical values of strings $\sti{x}$, and length of a string $|x|$, for all string variables $x$ in $\vars$.
%Formally, assume that $\vars=\{x_i,\ldots,x_n\}$ and $\varn=\{y_i,\ldots,y_m\}$, a length constraint is of the form $(\sum_{i\in[1,n]}(j_i \times \sti{x_i}+k_i\times |x_i|)+ \sum_{i\in[1,m]}(l_i \times y_i)) \odot k$, where $\odot \in \{>,\geq, =, \leq, <\}$, and $j_i,k_i,l_i,k\in \mathbb{Z}$. We define $\modelof{\phi}= \{I \mid (\sum_{i\in[1,n]}(j_i \times \sti{I(x_i)}+k_i\times |I(x_i)|)+ \sum_{i\in[1,m]}(l_i \times I(y_i))) \odot k \}$.
We define  $\modelof{\phi}= \{I \mid I(\phi)= \true \}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Symbolic Flat Automata} \label{section:sfa}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Fix a finite set of character variables $\cvar$ ranging over $\Sigma_\epsilon$.
We call a string over $\cvar$ a symbolic string.
A interpretation $I$ over $\cvar$ is a mapping from $\cvar$ to $\Sigma_\epsilon$.
For a symbolic string $x= v_1v_2\cdots v_k$ over $\cvar$, $I(x)$ is defined as $I(v_1)\cdot I(v_2)\cdot \ldots \cdot I(v_k)$.

We introduce \emph{symbolic flat automata} (SFA) $(Q,T,\cvar_\epsilon,q_0,q_m)$ that will be used to define patterns for the variable domain restriction. SFA is a finite-state automaton operating over the alphabet $\cvar_\epsilon$ and with additional structural constraints.
\begin{enumerate}
	\item the initial state $q_0$ and final state $q_m$ are connected by a straight-line structure, i.e., $(q_i,v_i,q_i+1) \in T$ and states $q_i \in Q$ for $i\in[0,m-1]$.
	\item  each state $q_i$ may be included to an unique simple cycle, i.e. $q_i^j \in Q$ and $(q_i, v_i^0, q_i^1), (q_j, v_i^j, q_i^{j+1}), (q_i^l, v_i^l, q_i) \in T$ for $i\in [0,n]$ and $j\in [1,l-1]$. 
	\item each variable in $\cvar$ appears in at most one transition (but $\epsilon$ transition can appear multiple times). 
\end{enumerate} 
Such automata are recognizers of symbolic strings, i.e., strings over $\cvar$.

\begin{figure}
	\tikzset{state/.style={circle,draw=blue!50,fill=blue!20,
			thick,inner sep=0pt,minimum size=6mm}, initial text=$ $}
	\begin{tikzpicture} 
		\node[state,initial] (q0)  {$q_0$};
		\node[state] (q1) [right = 3cm of q0] {$q_1$};
		\node (q2) [right = 1.8cm of q1]{};
		\node (q3) [right = 1cm of q2]{};
		\node[state,accepting] (qm) [right = 1.8cm of q3] {$q_m$};

		\node[state] (q01) [above left = 0.6cm of q0] {$q_0^1$};
		\node[state] (q02) [above = 1cm of q0] {$q_0^2$};
		\node[state] (q03) [above right = 0.6cm of q0] {$q_0^3$};

		\node[state] (q11) [above left = 0.8cm of q1] {$q_1^1$};
		\node[state] (q12) [above right = 0.8cm of q1] {$q_1^2$};

		\node[state] (qm1) [above left = 0.6cm of qm] {$q_m^1$};
		\node[state] (qm2) [above = 1cm of qm] {$q_m^2$};
		\node[state] (qm3) [above right = 0.6cm of qm] {$q_m^3$};

  		\draw[->] (q0) edge [bend left] node [left]{$v_0^0$} (q01) ;
  		\draw[->] (q01) edge [bend left] node [left]{$v_0^1$} (q02) ;
  		\draw[->] (q02) edge [bend left] node [right]{$v_0^2$} (q03) ;
  		\draw[->] (q03) edge [bend left] node [right]{$v_0^3$} (q0) ;

@@

  		\draw[->] (qm) edge [bend left] node [left]{$v_m^0$} (qm1) ;
		\draw[->] (qm1) edge [bend left] node [left]{$v_m^1$} (qm2) ;
		\draw[->] (qm2) edge [bend left] node [right]{$v_m^2$} (qm3) ;
		\draw[->] (qm3) edge [bend left] node [right]{$v_m^3$} (qm) ;

  		\draw[->] (q1) edge [bend left] node [left]{$v_1^0$} (q11) ;
		\draw[->] (q11) edge [bend left] node [above]{$v_1^1$} (q12) ;
		\draw[->] (q12) edge [bend left] node [right]{$v_1^2$} (q1) ;
	\end{tikzpicture} 
	\label{fig:sfa_def}
	\caption{An example of a symbolic flat automaton}
\end{figure}

A FA $D=(Q,T',\Sigma_\epsilon,q_0,q_m)$ is an \emph{instance} of a SFA $S=(Q,T,\cvar_\epsilon,q_0,q_m)$, if there exists an interpretation $I: \cvar_\epsilon \rightarrow \Sigma_\epsilon$ satisfying $I(\epsilon)=\epsilon$ and $T'=\{(q,I(c),q')\mid (q,c,q')\in T\}$. We write $I(S)$ to denote the instance of $S$ wrt. $I$. The language of $S$ is defined as $L(S)= \cup\{L(D) \mid D \mbox{ is an instance of } S\}$. Notice that $L(S)$ is regular because every SFA has only finitely many instance and the finite union of regular languages are also regular.

For such kind of automata, the number of occurrence of each symbol in $\cvar$ uniquely characterize an accepting string. For example, in the example of Figure~\ref{fig:sfa_def}, if the string $v_0(v^0_1v^1_1v^2_1)^2v_1v_{m-1}$ is accepted, then all other strings $w$ with $|w|_{v_0}=|w|_{v_1}=\ldots=|w|_{v_{m-1}}=1$ and also $|w|_{v^0_1}=|w|_{v^1_1}=|w|_{v^2_1}=2$ are not accepted. Moreover, from $|w|_{v_0}=|w|_{v_1}=\ldots=|w|_{v_{m-1}}=1$ and $|w|_{v^0_1}=|w|_{v^1_1}=|w|_{v^2_1}=2$, we can derive $w=v_0(v^0_1v^1_1v^2_1)^2v_1v_{m-1}$ by exploring the automata structure. \todo{add Presburger constraints to allow more expressive power.}

@@

Assume that $S_n=(Q,T,\cvar^1_\epsilon,q_i,q_f)$ is the SFA of $x_1\cdot x_2 \cdots x_n$ and $S_m=(P,T',\cvar^2_\epsilon,p_i,p_f)$ is the SFA of $x_{n+1}\cdot x_{n+2} \cdots x_m$. Our next task is to find a model for $x_1\cdot x_2 \cdots x_n = x_{n+1}\cdot x_{n+2} \cdots x_m$ under the domain restriction specified by $S_n$ and $S_m$.

This is equivalent to finding a pair of symbolic strings $(w_n,w_m) \in L(S_n)\times L(S_m)$ and a common interpretation $I$ such that 
\begin{enumerate}
	\item $I(w_n)=I(w_m)$
	\item $I$ assign to variants of the same variable the same value. For example, in Figure~\ref{fig:sfa} (c) and (d), $v_0^{(0)}$ and $v_0^{(1)}$ should be assigned the same value.
	\item variants of the same variable should occurs equally often in $w_n\cdot w_m$. For example, in Figure~\ref{fig:sfa} (c) and (d), $v_0^{(0)}v_1^{(0)}v_0^{(0)}v_1^{(0)}$ and $v_0^{(1)}v_1^{(1)}v_0^{(1)}v_1^{(1)}$ is a pair of symbolic strings satisfying this condition. The number of occurrences of $v_i^{(0)}$ is the same to $v_i^{(1)}$ for $i\in[0,1]$ in the concatenation of the two symbolic strings .
	But the pair $v_0^{(0)}v_1^{(0)}v_0^{(0)}v_1^{(0)}$ and $v_2^{(0)}v_3^{(0)}v_2^{(0)}v_3^{(0)}$ does not satisfy the condition. Observe that $v_0^{(0)}$ occurs twice, but $v_0^{(1)}$ is not there.
\end{enumerate}

@@

Such product automaton uses $Q\times P$ as the set of states and $\cvar^1_\epsilon \times \cvar^2_\epsilon$ as the alphabet. If a transition $((q_1,p_1), (v_i^{(j)},v_{i'}^{(j')}),(q_2,p_2))$ is taken, for some $i,j,i',j' \in \mathbb{N}$, it means the character variable $v_i$ and $v_{i'}$ (and so all of their indexed variants) should be assigned the same value and this value leads $S_n$ from state $q_1$ to state $q_2$ and $S_m$ from $p_1$ to $p_2$.

More precisely, the product automaton is a tuple $(Q\times P, T^2, \cvar^1_\epsilon \times \cvar^2_\epsilon, (q_i,p_i),(q_f,p_f))$, where the transition relation $T^2$ is the minimal set satisfying the following.

>>>>

\title{Handling String-Number Conversion in String Constraint Solving}
\author{}
\institute{}

@@

\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\sti}[1]{\mbox{\textbf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textbf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvar}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{\mbox{$\mathbb{V}^1_{\Sigma_\epsilon}$}}
\newcommand{\cvartwo}{\mbox{$\mathbb{V}^2_{\Sigma_\epsilon}$}}
\newcommand{\modelof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}

@@

\todo{Why this problem is important? Need to say it is used often, e.g., read 
from text file or user input. Why this problem is challenging? It is undecidable. Some tricky cases 
using string as array index. Bounded SAT-based encoding might not work even for 
very simple constraints. e.g., $y=str2int(x) \wedge y>9999999$, if $x$ is a 
ASCII string, then XXX Boolean variables are required. Or we can just propose 
an example that all major solver fails. Maybe we still say our approach is still over-approximation + under-approximation, in this case, need to think how to present the over-approximation part. Maybe no CEGAR is fine.}

In this paper, we propose an approach that handles string constraints with string-number conversion efficiently. This is a very challenging problem. From the theoretical point of view, this problem is already proven to be undecidable~\cite{vijay_rp}. In practice, in our experiments, all of the state-of-the-art string constraint solvers do not support string-number conversion correctly. Some of them handles such constraints with only over-approximation of the original formula, and is known to have one-side error (see Section~\ref{section:experiment} for detailed explanation). Here we propose a framework that uses both over and under-approximation of the string constraints. The over-approximation is for proving UNSAT and the under-approximation is for proving SAT of the string constraint. Both over and under-approximation should fail in a decidable fragment of string constraints that can be efficiently solved.

For over-approximation, there are a few existing algorithms~\cite{atva2019,z3}, our framework does not set any restriction on which algorithm to be used. In our experiences, the UNSAT instances in all available benchmarks are not that difficult and many approaches can handle all of them almost perfectly. The battle field is in fact those SAT instances. 
In our prototype tool, we implement the algorithm presented in~\cite{atva2019}, in addition with some additional axioms, to handle the over-approximation of string constraints with number-string conversion.

For under-approximation, we restrict the search space of each string variable to strings that obey some predefined and parameterized pattern. By adjusting the parameters, one can easily enlarge or prune out the potential solution space. In paper, we propose to use patterns defined by \emph{symbolic flat automata} (SFA), the automata are very expressive yet allows very efficiently manipulation. For example, it avoids the alphabet 
explosion problem that the approach in~\cite{PLDI2017} suffers.

Based on the SFA encoding, we manage to translate the string constraint with string-number function to an equisatisfiable pure numerical constraints. Thus the problem is significantly simplified. We show that if we restrict the variable to arbitrary symbolic flat automata, one can convert the a string constraint with string-number function to a integer constraint consisting of both polynomials and exponentials (formula~\ref{formula:str_int}). To the best of our knowledge, the satisfiability prolem of integer constraints with a mix of polynomials and exponentials is still open.

For the particular case with only one integer variable, the satisfiability can be decided using the algorithm in~\cite{POPL2019}. However, the algorithm involves a quantifier elimination procedure, which is double-exponential to the length of the input formula and hence cannot handle large instances.

For efficiency reason, we propose to use \emph{SFA with only one loop} to handle string-number conversion function. That is, if we encountered a constraint $y=\sti{x}$, then we always restrict the domain of $x$ to some SFA with one loop, instead of arbitrary SFAs. This allows us to convert a string constraint with string-number conversion function to a equisatisfiable linear integer constraint, which can be solved very efficiently.

\todo{Talk about experiments}

To ease presentation, in the main text, we only consider string constraints consisting of \emph{equality constraints (a.k.a. word equation)}, \emph{length constraints}, and \emph{string-number conversion} functions. We describe the extension to support transducer and membership constraints in the appendix.

@@

A \emph{finite automaton} is a tuple $(Q,T,\Sigma,q_0,q_m)$, where $Q$ is the set of states, $T\subseteq Q\times (\Sigma\cup \{\epsilon\} )\times Q $ is the set of transition relation, $\Sigma$ is the alphabet, $q_0$ is the initial state, and $q_m$ is the final state. The semantics of finite automaton is defined in the standard manner. We use $L(A)$ to denote the regular language, i.e., set of accepted strings, of a finite automaton $A$.

In this paper, we fix a finite alphabet $\Sigma \subseteq \mathbb{N}$. Note that here we assume the alphabet is a finite subset of natural numbers. Essentially, we try to capture the numerical encoding of the corresponding symbols in computers (e.g., in ASCII, `A' is encoded as $65$). We assume there is a one-one mapping between numbers in $\Sigma$ and the character it encodes. For simplicity of presentation, we assume the character `$0$' is mapped to number $0$, `$1$' to $1$,$\ldots$, and `$9$' to $9$. For other character $c$, we use $\enc{c}$ to denote the number that it maps to. Notice that this approach is general enough to support any finite set of characters. We write $[n,m]$ for the set of numbers (or symbols in $\Sigma$) $\{k\mid n\leq k \leq m\}$. A \emph{term} is an element in $(\vars\cup \Sigma)^*$.

Given a finite automaton $A$ over $\{a_1,a_2,\ldots,a_n\}$, the Parikh image of $L$ is the set $\{(|w|_{a_1},\ldots, |w|_{a_n}) \mid w \in  L(A) \}$. Moreover, one can construct from $A$ in linear time a Presburger formula $\phi_A(v_1, \ldots,v_n)$ such that $\phi(c_1, \ldots,c_n) \iff (c_1, \ldots,c_n)$ in the Parikh image of $L$~\cite{peter}.

We use $\vars$ to denote a finite set of string variables ranging over $\Sigma^*$, $\varn$ to denote a finite set of integer variables ranging over $\mathbb{Z}$, and $\cvar$ to denote a finite set of character variables  ranging over $\Sigma_\epsilon$. Sometimes we use index to denote different set of variables over the same range, e.g., $\cvarone$ and $\cvartwo$. 
We call a string over $\cvar$ a \emph{symbolic string}. 
For a variable $x$, we call its primed version (e.g., $x'$) or indexed versions (e.g., $x_1$, $x^1$) its \emph{variants}.  
In this paper, we use $x,y$ or their variants to denote variables in $\vars$, $n$ or its variants to denote variables in $\varn$, and $v$ and its variants for variables in $\cvar$.

An \emph{interpretation over $\vars$ and $\varn$} is a mapping $I$ from $\vars\cup \varn$ to $\Sigma^* \cup \mathbb{N}$. We lift $i$ to terms by defining $I: (\vars\cup \Sigma)^* \rightarrow \Sigma^*$ with $I(a) =a$ for $a\in \Sigma$ and $I(t_1 \cdot t_2) = I(t_1)\cdot I(t_2)$.
An \emph{interpretation $I$ over $\cvar$} is a mapping from $\cvar$ to $\Sigma_\epsilon$. For a symbolic string $x= v_1v_2\cdots v_k$ over $\cvar$, $I(x)$ is defined as $I(v_1)\cdot I(v_2)\cdot \ldots \cdot I(v_k)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{String Constraints} \label{section:sc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \emph{string-number conversion function} $\sti{x}$ is defined as follows. For $a\in [0,9]$, we have $\sti{a}=a$ and for $x \cdot a \in [0,9]^+$, $\sti{x\cdot a} = 10\times \sti{x}+a$. For $x\notin [0,9]^+$, $\sti{x}=-1$.

Notice that only positive integer is supported in this semantics. This is the semantics used by most of the SMT solvers, and hence we follow it in this paper. The encoding has a benefit that only can also handle the case where $x$ is ``not a number'', using the condition $sti{x} = -1$.

Supporting only positive integer is not a strong restriction, since conversing from negative integer can still be encoded using the positive only version. More specifically, if we want to say $n \in \varn$ is the integer value of the string $x \in \Sigma^*$, we can write the following constraint (ignore the case that $x$ is ``not a number'', which can also be encoded using more conditions):

$$(n= \sti{x}) \vee (n=-\sti{x'} \wedge x = \enc{-}\cdot x')$$
Recall that $\enc{-}$ is the integer encoding of the minus symbol `$-$'. The \emph{number-string conversion function} $\its{n}$ is defined symmetrically. We have $x = \its{n}$ iff $y= \sti{n}$.

@@

An \emph{integer constraint} $\phi$ is a linear constraint over the variables in $\varn$ and length of a string $|x|$ for all $x \in \vars$.
%Formally, assume that $\vars=\{x_i,\ldots,x_n\}$ and $\varn=\{y_i,\ldots,y_m\}$, a length constraint is of the form $(\sum_{i\in[1,n]}(j_i \times \sti{x_i}+k_i\times |x_i|)+ \sum_{i\in[1,m]}(l_i \times y_i)) \odot k$, where $\odot \in \{>,\geq, =, \leq, <\}$, and $j_i,k_i,l_i,k\in \mathbb{Z}$. We define $\modelof{\phi}= \{I \mid (\sum_{i\in[1,n]}(j_i \times \sti{I(x_i)}+k_i\times |I(x_i)|)+ \sum_{i\in[1,m]}(l_i \times I(y_i))) \odot k \}$.
We define  $\modelof{\phi}= \{I \mid I(\phi)= \true \}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Over-Approximation and Under-Approximation Framework} \label{section:framework}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Symbolic Flat Automata} \label{section:sfa}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\st}[2]{q_{#1}^{#2}}
\newcommand{\sym}[2]{v_{#1}^{#2}}

We introduce \emph{symbolic flat automata} (SFA) $(Q,T,\cvar,\st{0}{0},\st{m}{0})$ that will be used to define patterns for the string variable domain restriction. SFA is a finite automaton operating over the alphabet $\cvar$ in addition with the following structural constraints.
\begin{enumerate}
	\item the initial state $\st{0}{0}$ and final state $\st{m}{0}$ are connected by a straight-line structure, i.e., $(\st{i}{0},v_i,\st{i+1}{0}) \in T$ and states $\st{i}{0} \in Q$ for $i\in[0,m-1]$.
	\item  each state $\st{i}{0}$ may be included to an unique simple cycle. Assume that $\st{i}{0}$ is included in a cycle with $l_i$ states (includes $\st{i}{0}$), then we have $\st{i}{j} \in Q$ and $(\st{i}{j-1}, \sym{i}{j-1}, \st{i}{j \bmod l_i})$, for $j\in [1,l_i]$. 
	\item each variable in $\cvar$ appears in at most one transition (but $\epsilon$ transition can appear multiple times). 
\end{enumerate} 
Such automata are recognizers of symbolic strings, i.e., strings over $\cvar$.

\begin{figure}
	\tikzset{state/.style={circle,draw=blue!50,fill=blue!20,
			thick,inner sep=0pt,minimum size=6mm}, initial text=$ $}
	\begin{tikzpicture} 
		\node[state,initial] (q0)  {$\st{0}{0}$};
		\node[state] (q1) [right = 3cm of q0] {$\st{1}{0}$};
		\node (q2) [right = 1.8cm of q1]{};
		\node (q3) [right = 1cm of q2]{};
		\node[state,accepting] (qm) [right = 1.8cm of q3] {$\st{m}{0}$};

		\node[state] (q01) [above left = 0.6cm of q0] {$\st{0}{1}$};
		\node[state] (q02) [above = 1cm of q0] {$\st{0}{2}$};
		\node[state] (q03) [above right = 0.6cm of q0] {$\st{0}{3}$};

		\node[state] (q11) [above left = 0.8cm of q1] {$\st{1}{1}$};
		\node[state] (q12) [above right = 0.8cm of q1] {$\st{1}{2}$};

		\node[state] (qm1) [above left = 0.6cm of qm] {$\st{m}{1}$};
		\node[state] (qm2) [above = 1cm of qm] {$\st{m}{2}$};
		\node[state] (qm3) [above right = 0.6cm of qm] {$\st{m}{3}$};

  		\draw[->] (q0) edge [bend left] node [left]{$\sym{0}{0}$} (q01) ;
  		\draw[->] (q01) edge [bend left] node [left]{$\sym{0}{1}$} (q02) ;
  		\draw[->] (q02) edge [bend left] node [right]{$\sym{0}{2}$} (q03) ;
  		\draw[->] (q03) edge [bend left] node [right]{$\sym{0}{3}$} (q0) ;

@@

  		\draw[->] (qm) edge [bend left] node [left]{$\sym{m}{0}$} (qm1) ;
		\draw[->] (qm1) edge [bend left] node [left]{$\sym{m}{1}$} (qm2) ;
		\draw[->] (qm2) edge [bend left] node [right]{$\sym{m}{2}$} (qm3) ;
		\draw[->] (qm3) edge [bend left] node [right]{$\sym{m}{3}$} (qm) ;

  		\draw[->] (q1) edge [bend left] node [left]{$\sym{1}{0}$} (q11) ;
		\draw[->] (q11) edge [bend left] node [above]{$\sym{1}{1}$} (q12) ;
		\draw[->] (q12) edge [bend left] node [right]{$\sym{1}{2}$} (q1) ;
	\end{tikzpicture} 
	\label{fig:sfa_def}
	\caption{An example of a symbolic flat automaton}
\end{figure}

A finite automaton $D=(Q,T',\Sigma,q_0,q_m)$ is an \emph{instance} of a SFA $S=(Q,T,\cvar,q_0,q_m)$, if there exists an interpretation $I: \cvar \cup \{\epsilon\} \rightarrow \Sigma_\epsilon$ satisfying $I(\epsilon)=\epsilon$ and $T'=\{(q,I(c),q')\mid (q,c,q')\in T\}$. We write $I(S)$ to denote the instance of $S$ wrt. $I$. The language of $S$ is defined as $L(S)= \cup\{L(D) \mid D \mbox{ is an instance of } S\}$. Notice that $L(S)$ is regular because every SFA has only finitely many instance and the finite union of regular languages are also regular.

For such kind of automata, the number of occurrence of each symbol in $\cvar$ uniquely characterize an accepting symbolic string. For example, in the example of Figure~\ref{fig:sfa_def}, if the string $v_0(v^0_1v^1_1v^2_1)^2v_1v_{m-1}$ is accepted, then all other strings $w$ with $|w|_{v_0}=|w|_{v_1}=\ldots=|w|_{v_{m-1}}=1$ and also $|w|_{v^0_1}=|w|_{v^1_1}=|w|_{v^2_1}=2$ are not accepted. Moreover, from $|w|_{v_0}=|w|_{v_1}=\ldots=|w|_{v_{m-1}}=1$ and $|w|_{v^0_1}=|w|_{v^1_1}=|w|_{v^2_1}=2$, we can derive $w=v_0(v^0_1v^1_1v^2_1)^2v_1v_{m-1}$ by exploring the automata structure. \todo{add Presburger constraints to allow more expressive power.}

@@

Assume that $S_n=(Q,T,\cvarone,q_i,q_f)$ is the SFA of $x_1\cdot x_2 \cdots x_n$ and $S_m=(P,T',\cvartwo,p_i,p_f)$ is the SFA of $x_{n+1}\cdot x_{n+2} \cdots x_m$. Our next task is to find a model for $x_1\cdot x_2 \cdots x_n = x_{n+1}\cdot x_{n+2} \cdots x_m$ under the domain restriction specified by $S_n$ and $S_m$. 
Note that $\cvarone \cap \cvartwo =\emptyset$, because all variables will be renamed to some fresh variable.

This is equivalent to finding a pair of symbolic strings $(w_n,w_m) \in L(S_n)\times L(S_m)$ and a common interpretation $I$ such that 
\begin{enumerate}
	\item $I(w_n)=I(w_m)$
	\item $I$ assign to variants of the same variable the same value. For example, in Figure~\ref{fig:sfa} (c) and (d), $v_0^{(0)}$ and $v_0^{(1)}$ should be assigned the same value.
	\item different versions of the same variable should occurs equally often in $w_n\cdot w_m$. For example, in Figure~\ref{fig:sfa} (c) and (d), $w_{xy}=v_0^{(0)}v_1^{(0)}v_0^{(0)}v_1^{(0)}$ and $w_{yx}=v_0^{(1)}v_1^{(1)}v_0^{(1)}v_1^{(1)}$ is a pair of symbolic strings satisfying this condition. The number of occurrences of $v_i^{(0)}$ is the same to $v_i^{(1)}$ for $i\in[0,1]$ in the concatenation $w_{xy}\cdot w_{yx}$.
	But the pair $v_0^{(0)}v_1^{(0)}v_0^{(0)}v_1^{(0)}$ and $v_2^{(0)}v_3^{(0)}v_2^{(0)}v_3^{(0)}$ does not satisfy the condition. Observe that $v_0^{(0)}$ occurs twice, but $v_0^{(1)}$ is not there.
\end{enumerate}

@@

Such product automaton uses $Q\times P$ as the set of states and $(\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} )$ as the alphabet. If a transition $((q_1,p_1), (v_i^{(j)},v_{i'}^{(j')}),(q_2,p_2))$ is taken, for some $i,j,i',j' \in \mathbb{N}$, it means the character variable $v_i$ and $v_{i'}$ (and so all of their indexed variants) should be assigned the same value and this value leads $S_n$ from state $q_1$ to state $q_2$ and $S_m$ from $p_1$ to $p_2$.

More precisely, the product automaton is a tuple $(Q\times P, T^2, (\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} ), (q_i,p_i),(q_f,p_f))$, where the transition relation $T^2$ is the minimal set satisfying the following.
